---
title: "Comment Mining Write Up"
author: "Anders Swanson"
output: html_document
bibliography: library.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### **Research Question**  

#### Ideal question
> What effect do teacher-written trimester comments have on future student performance?*

#### Realistic question
> How does the language used to describe students in trimester one comments vary depending on level of improvement in trimester two?

### **Purpose**  

##### 1. Seek out what an "effective" comment is.  
Full disclosure:  
* I strongly dislike writing comments.  
* I think WAY too much.  

I just finished four years at Xiamen International School: a private, MYP K-12 school with three "grading" periods a year. At the end of every trimester the system teachers reported student's achievement against the MYP subject-specific criteria rubrics (e.g. in Mathematics: A: Knowing and Understanding, B: Investigating Patterns, C: Communicating, D: Applying Math in Real-Life) that were used that trimester.

In addition to criterion-based reporting, teachers wrote a three-sentence narrative comment about the students achievement in this format:  
1. An area in which the student did well  
2. An area in which the student could improve  
3. A recommendation as to how the student could improve in the above area.  


 > The report card should, above all else, be user-friendly: Parents must be able to easily understand the information it contains. --*Grant Wiggins* [@Wiggins1994]

##### 2. Enhance Creativity of Teachers  
I love the conversations I've heard in workshops concerning "data-driven" learning and action research. However, the majority of tools for educational data analysis (MAP, SAT, Atlas Rubicon) the intended audience is really the administrators. I want to focus on practice-based recommendations for helping teachers teach.

##### 3. Learn R
I started this project having only competed the first few courses of the Johns Hopkins Data Science specialization series on Coursera. I knew basics of R but I hoped to become more fluent as I went. I've definitely learned a lot and moved beyond the syntax to understand more about R, especially packages like tm, dplyr, and ggplot2.


***
### **Principles**  
##### 1. Share results with community  

This article is a start. [Full GitHub repository is here](https://github.com/swanderz/xis_mining)

##### 2. Produce reliable, readable, commented code  

This is something I've I'd love to share the data set but it contains and scrubbing all names from comments is something beyond my difficultly level currently. 

"data can either be useful or perfectly anonymous but never both."

##### 3. Be as objective as possible  
I started this project having only competed the first few courses of the Johns Hopkins Data Science specialization series on Coursera. I knew basics of R but I hoped to become more fluent as I went.  

##### 4. Strive for "good enough"  
I've started and stopped this project three times now over the course of a year. I've branched off and explored MANY different aspects of this dataset but in the end the story is original


### EDA - Exploratory Data Analysis  

First I'll start by loading all the extra packages I used for my analysis.

```{r libraries & source, message = FALSE}
#load required libraries, data, and created functions
library(tm)
library(dplyr)
library(tidyr)
library(RWeka)
library(xlsx)
library(lubridate)
library(ggplot2)
library(stringr)
library(devtools)
library(knitr)

#set wd and paths and source functions
setwd("/Users/andersswanson/Desktop/comment\ mining")
source("Functions.R")
source("MB - t12 growth.R")
```

Then I'll load the csv file I got by exporting term grades from ManageBac
```{r t1.report, , warning = FALSE}
 t1.report <- GetReportsDFfromMBcsv("data/t1 comments.csv")

#find dimensions (rows & columns) of the table
dim(t1.report)

#get column names
colnames(t1.report)
```

So this table has 14 columns 1066 rows, where each row is one student's report for one class

```{r sample table, warning = FALSE}


t1.report.ex <- t1.report %>%
        select(-Student.Comment) %>%
        sample_n(8)
        #head(10)
t1.report.ex$Student.ID <- "1000****"
t1.report.ex$First.Name <- "****"
t1.report.ex$Last.Name <- "****"
kable(t1.report.ex, digits = 2)


```

The above table shows eight randomly selected reports. 
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# References  
