---
title: "Comment Mining Write Up"
author: "Anders Swanson"
output: html_document
bibliography: library.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### **Purpose**  

##### 1. Seek out what an "effective" comment is.  
Full disclosure:  
* I strongly dislike writing comments.  
* I think WAY too much.  

I just finished four years at Xiamen International School: a private, MYP K-12 school with three "grading" periods a year. At the end of every trimester the system teachers reported student's achievement against the MYP subject-specific criteria rubrics (e.g. in Mathematics: A: Knowing and Understanding, B: Investigating Patterns, C: Communicating, D: Applying Math in Real-Life) that were used that trimester.

In addition to criterion-based reporting, teachers wrote a three-sentence narrative comment about the students achievement in this format:  
1. An area in which the student did well  
2. An area in which the student could improve  
3. A recommendation as to how the student could improve in the above area.  


 > The report card should, above all else, be user-friendly: Parents must be able to easily understand the information it contains. --*Ender Wiggins* [@Wiggins1994]

##### 2. Enhance Creativity of Teachers  
I love the conversations I've heard in workshops concerning "data-driven" learning and action research. However, the majority of tools for educational data analysis (MAP, SAT, Atlas Rubicon) the intended audience is really the adminstrators. 

##### 3. Learn R
I started this project having only competed the first few courses of the Johns Hopkins Data Science specialization series on Coursera. I knew basics of R but I hoped to become more fluent as I went. 


***
### **Principles**  
##### 1. Share results with community  

This article is a start. [Full GitHub repository is here](https://github.com/swanderz/xis_mining)

##### 2. Produce reliable, readable, commented code  

This is something I've I'd love to share the dataset but it contains and scrubbing all names from comments is something beyond my difficultly level currently. 

"data can either be useful or perfectly anonymous but never both."

##### 3. Be as objective as possible  
I started this project having only competed the first few courses of the Johns Hopkins Data Science specialization series on Coursera. I knew basics of R but I hoped to become more fluent as I went.  

##### 4. Strive for "good enough"  

***
### EDA - Exploratory Data Analysis  

I'd like to do a basic repository

```{r, eval = FALSE }
#load required libraries, data, and created functions
library(tm)
library(dplyr)
library(tidyr)
library(RWeka)
library(xlsx)
library(lubridate)
library(ggplot2)
library(stringr)
library(devtools)
```


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# References  
