---
title: "Comment Mining Write Up"
author: "Anders Swanson"
output: html_document
bibliography: library.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### PURPOSE  
#### *(i.e. By completing this project, I would like to...)*  


##### 1. Seek out what an "effective" comment is.  
Full disclosure:  
* I strongly dislike writing comments.  
* I think WAY too much.  

I just finished four years at Xiamen International School: a private, MYP K-12 school with three "grading" periods a year.

 > The report card should, above all else, be user-friendly: Parents must be able to easily understand the information it contains. --*Ender Wiggins* [@Wiggins1994]

##### 2. Enhance Creativity of Teachers  
I love the conversations I've heard in workshops concerning "data-driven" learning and action research. However, the majority of tools for educational data analysis (MAP, SAT, Atlas Rubicon) the intended audience is really the adminstrators. I'm very  

##### 3. Learn R
I started this project having only competed the first few courses of the Johns Hopkins Data Science specialization series on Coursera. I knew basics of R but I hoped to become more fluent as I went. 


---
## Principles
### 1. Share results with community  
That's why I've got the notebook!  

### 2. Produce reliable, readable, commented code  
I love the conversations I've heard in workshops concerning "data-driven" learning and action research. However, the majority of tools for educational data analysis (MAP, SAT, Atlas Rubicon) the intended audience is really the adminstrators. I'm very  

### 3. Be as objective as possible  
I started this project having only competed the first few courses of the Johns Hopkins Data Science specialization series on Coursera. I knew basics of R but I hoped to become more fluent as I went.  

### 4. Strive for "good enough"  


```{r }

```


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# References  
